---
title: "*lac* induction and growth lag in naive cells"
author: Thomas Julou
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

## Distribution of lags

For the first switch, all conditions (without IPTG) should be equivalent:
  
```{r}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_date) %>%
  filter(switch_idx==1) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
  mutate(lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         lag=ifelse(is.na(lag), -1, lag)) %>% 
  ggplot(aes(x=lag/60)) +
  stat_ecdf(aes(y=1-..y.., col=variable)) +
  labs(x='lag after the switch (min)', y='reverse cumulative probability', title='distribution of lags at 1st switch (without 8h)') +
  expand_limits(x=0)

mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_date) %>%
  filter(switch_idx==1) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
  mutate(lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         lag=ifelse(is.na(lag), -200, lag)) %>% 
  ggplot(aes(x=lag/60)) +
  facet_grid(variable~.) +
  geom_step_hist(aes(col=factor(switch_idx)), binwidth=6, position='identity') +
  geom_histogram(aes(fill=factor(switch_idx)), binwidth=6, position='identity', 
                 col='transparent', alpha=.2) +
  labs(x='lag after the switch (min)', title='distribution of finite lags at 1st switch (without 8h)', col='switch', fill='switch') +
  xlim(0, 240) +
  guides(col='none', fill='none')

```


```{r BZSposter, eval=FALSE}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_date) %>%
  filter(switch_idx==1) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  gather(variable, lag, growth_lag, gfp_lag, factor_key=TRUE) %>% 
  mutate(variable=forcats::fct_recode(variable, 'growth lag'='growth_lag', 'lac induction lag'='gfp_lag'),
         lag=ifelse(lag>240*60, Inf, lag) ) %>% #, 
         # lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         # lag=ifelse(is.na(lag), -200, lag)) %>% 
  ggplot(aes(x=lag/60)) +
  # facet_grid(variable~.) +
  geom_histogram(aes(fill=variable), binwidth=6, position='identity', 
                 col='transparent', alpha=.1) +
  geom_step_hist(aes(col=variable), binwidth=6, position='identity') +
  labs(x='lag (min)') +
      scale_x_continuous(breaks=seq(0, 200, 50)) +
      theme_classic(base_size=8) + theme(axis.line.x=element_line(), axis.line.y=element_line()) +
      theme(legend.position=c(1, 1), legend.justification=c(1, 1) )
ggsave('plots/BZSposter2017/1st_switch_lags.pdf', width=180/25.4/2, height=100/25.4/2)
  
```


how the bimodal lag of GFP induction becomes more unimodal for growth is not clear: it is tempting t hypothesize that there is some additional variability in the time different cells need to restart growing after expressing GFP / that they induce GFP at different rates, hence bluring the 2 peaks into a single broader peak.

```{r}
mycells_switching %>% 
  mutate(lag_200=as.character(lag_200), lag_200=ifelse(is.na(lag_200), 'NA', lag_200),
         lag_200=ifelse(is.finite(as.numeric(lag_200)) & as.numeric(lag_200)<50*60, '<50', lag_200),
         lag_200=ifelse(is.finite(as.numeric(lag_200)) & as.numeric(lag_200)>=50*60, '≥50', lag_200) ) %>%
  group_by(switch_idx, lag_200) %>% 
  summarise(n=n()) %>% 
  group_by(switch_idx) %>% mutate(p=n/sum(n)) %>% 
  knitr::kable(digits=3)

```

Check the fraction of arrested cells in mother cells since they have no bias of long lags being more truncated:

```{r}
mycells_switching %>% 
  filter(cell_num_from_bottom==0) %>% # keep only mother cells
  mutate(lag_200=as.character(lag_200), lag_200=ifelse(is.na(lag_200), 'NA', lag_200),
         lag_200=ifelse(is.finite(as.numeric(lag_200)) & as.numeric(lag_200)<50*60, '<50', lag_200),
         lag_200=ifelse(is.finite(as.numeric(lag_200)) & as.numeric(lag_200)>=50*60, '≥50', lag_200) ) %>%
  group_by(switch_idx, lag_200) %>% 
  summarise(n=n()) %>% 
  group_by(switch_idx) %>% mutate(p=n/sum(n)) %>% 
  knitr::kable(digits=3)

```



## Lag heritability

Let's look at the heritability of the lag in naive cells, i.e. how much a common history (and hence cellular state) determines the duration of the lag.
first visualize the pairwise correlations to understand the heritability estimator.

```{r}
filter(mypairs_switching, rel %in% c('sisters', 'cousins2')) %>% 
  filter(!date %in% discarded_date) %>%
  filter(switch_idx==1) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  mutate(value_1=ifelse(value_1<240*60, value_1, Inf), value_2=ifelse(value_2<240*60, value_2, Inf)) %>% 
  # filter(is.finite(value_1), is.finite(value_2)) %>% 
  ggplot(aes(value_1/60, value_2/60, col=rel)) +
# ggplot(filter(mypairs_switching, rel!='niece'), aes(lag_200_1/60, lag_200_2/60, col=rel)) +
  # facet_wrap(~switch_idx, scales='free', ncol = 3) +
     facet_grid(rel~variable) +
  geom_abline(col='gray50') +
  geom_point(alpha=.5, size=.5) +
  expand_limits(x=0, y=0) +
  labs(x='induction lag cell 1 (min)', y='induction lag cell 2 (min)', col='relationship') +
  scale_colour_manual(values=c('sisters'=brewer_cols[1], 'cousins2'=brewer_cols[3])) +
  guides(col='none') +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  coord_equal()

```

```{r}
filter(mypairs_switching, rel!='niece') %>% 
  filter(!date %in% discarded_date) %>%
  filter(switch_idx==1) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  mutate(value_1=ifelse(value_1<240*60, value_1, Inf), value_2=ifelse(value_2<240*60, value_2, Inf)) %>% 
  mutate(both_finite=is.finite(value_1) & is.finite(value_2)) %>%
  group_by(variable, rel) %>% 
  summarise(pearson=cor(value_1[both_finite], value_2[both_finite], method="pearson")^2, 
            spearman=cor(value_1, value_2, method="spearman", use="complete.obs")^2, 
            n=n(), n_finite=sum(both_finite)) %>% 
  gather(type, r2, pearson, spearman) %>% 
  # filter(type=='pearson') %>% 
  (function(.df) ggplot(.df, aes(rel, r2)) +
     facet_grid(type~variable, labeller=as_labeller(switch_facets_labels)) +
     geom_bar(aes(fill=rel), stat='identity', show.legend=FALSE) +
     geom_text(aes(y=0.99, label=paste0("n = ", n_finite)), angle=90, hjust=1) +
     expand_limits(y=1) + labs(x='relationship', y='squared pearson correlation r2') +
     theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1))
  )

```

By comparison, let's look at the heritability of various traits in constant environment:

```{r}
mypairs_constant %>% 
  filter(div_min==div_max) %>% 
  group_by(condition, variable, rel) %>% 
  summarise(r2=cor(value_1,  value_2, use="complete.obs", method="pearson")^2, n=n()) %>% 
  ggplot(aes(variable, r2)) +
  facet_wrap(~condition) +
  geom_bar(aes(fill=rel), stat='identity', pos='dodge') +
  geom_text(aes(y=1, label=sprintf("(n = %d)", n), group=rel), 
            angle=90, hjust=1, pos=position_dodge(width=0.8)) +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        legend.position='top')

```


```{r BZSposter, eval=FALSE}
bind_rows(
  mypairs_constant %>% 
  filter(div_min==div_max) %>% 
  filter(condition=='lactose', variable %in% c('alpha', 'l_birth', 'c_birth')) %>% 
  group_by(condition, variable, rel) %>% 
  summarise(r2=cor(value_1,  value_2, use="complete.obs", method="pearson")^2, n=n()),
  #
  filter(mypairs_switching, rel!='niece') %>% 
  filter(!date %in% discarded_date) %>%
  filter(switch_idx==1, variable %in% c('gfp_lag', 'growth_lag')) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  mutate(value_1=ifelse(value_1<240*60, value_1, Inf), value_2=ifelse(value_2<240*60, value_2, Inf)) %>% 
  mutate(both_finite=is.finite(value_1) & is.finite(value_2)) %>%
  group_by(variable, rel) %>% 
  summarise(r2=cor(value_1[both_finite], value_2[both_finite], method="pearson")^2, 
            n=sum(both_finite))
) %>% 
  ungroup %>% mutate(variable=factor(variable, 
    levels=c('gfp_lag', 'growth_lag', 'alpha', 'l_birth', 'c_birth')) %>% 
      forcats::fct_recode('lac induction\nlag'='gfp_lag', 'growth lag'='growth_lag', 'growth rate'='alpha', 'size at birth'='l_birth', '[GFP] at birth'='c_birth') ) %>% 
  ggplot(aes(variable, r2)) +
  geom_bar(aes(fill=rel), stat='identity', pos='dodge') +
  geom_text(aes(y=1, label=sprintf("(n = %d)", n), group=rel), 
            angle=90, hjust=1, vjust=.0, pos=position_dodge(width=0.8), size=2) +
  scale_x_discrete() +
  labs(y='Pearson correlation coeff r2', fill='relationship') +
  theme_classic(base_size=8) + 
  theme(axis.line.x=element_line(), axis.line.y=element_line()) +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        axis.title.x=element_blank())
ggsave('plots/BZSposter2017/heritability.pdf', width=180/25.4/2, height=100/25.4/2)

```

We want to understand (i) the molecular processes underlying the multimodal distribution of the lags, and (ii) what are theounds shared by cellullarcompurelated cells that determines their lags. Actually these questions might very well be linked. 


## Effect of spontaneous *lac* expression

hypothesized to explain how the positive feedback kicks off.
can be manipulated by titrating a given fraction of LacI with IPTG.

```{r}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_date) %>%
  filter(switch_idx==1) %>%
  filter(condition!='switch_lac_IPTG500uM') %>% 
  mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac',
         'lactose', condition)) %>%
  gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
  mutate(lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         lag=ifelse(is.na(lag), -1, lag)) %>% 
  ggplot(aes(x=lag/60)) +
  facet_grid(variable~.) +
  stat_ecdf(aes(y=1-..y.., col=condition)) +
  labs(x='lag after the switch (min)', y='reverse cumulative probability', title='distribution of lags at 1st switch') +
  expand_limits(x=0)

mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_date) %>%
  filter(switch_idx==1) %>%
  filter(condition!='switch_lac_IPTG500uM') %>% 
  mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac',
         'lactose', condition)) %>%
  gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
  mutate(lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         lag=ifelse(is.na(lag), -200, lag)) %>% 
  ggplot(aes(x=lag/60)) +
  facet_grid(variable~.) +
  geom_step_hist(aes(y=..density.., col=condition), binwidth=6, position='identity') +
  geom_histogram(aes(y=..density.., fill=condition), binwidth=6, position='identity', 
                 col='transparent', alpha=.2) +
  labs(x='lag after the switch (min)', title='distribution of finite lags at 1st switch') +
  xlim(0, 240)

```


Let's check that the *lac* promoter is not induced before the switch:

```{r}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_date) %>%
  filter(switch_idx==1) %>%
  filter(condition!='switch_lac_IPTG500uM') %>% 
  mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac',
         'lactose', condition)) %>%
  gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
  mutate(lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         lag=ifelse(is.na(lag), -200, lag)) %>% 
  ggplot(aes(condition, gfp_ini)) +
  geom_violin()
    # labs(x='lag after the switch (min)', title='distribution of finite lags at 1st switch') +
  # xlim(0, 240)

```


```{r BZSposter, eval=FALSE}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_date) %>%
  filter(switch_idx==1) %>%
  filter(condition!='switch_lac_IPTG500uM') %>% 
  mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac', 'no IPTG', condition),
         condition=forcats::fct_recode(condition, 'with 1µM IPTG'='switch_lac_withIPTG1uM',
                                       'with 5µM IPTG'='switch_lac_withIPTG5uM') ) %>%
  gather(variable, lag, growth_lag, gfp_lag, factor_key=TRUE) %>% 
  mutate(variable=forcats::fct_recode(variable, 'growth lag'='growth_lag', 'lac induction lag'='gfp_lag')) %>% 
  mutate(lag=ifelse(lag>240*60, Inf, lag),
         lag=ifelse(!is.infinite(lag), lag, 241*60),
         lag=ifelse(is.na(lag), -200, lag)) %>% 
  ggplot(aes(x=lag/60)) +
  facet_grid(variable~., scales='free_y') +
  geom_histogram(aes(y=..density.., fill=condition), binwidth=6, position='identity', 
                 col='transparent', alpha=.2) +
  geom_step_hist(aes(y=..density.., col=condition), binwidth=6, position='identity') +
  labs(x='lag (min)') +
  xlim(0, 240) +
  theme_classic(base_size=8) + 
  theme(axis.line.x=element_line(), axis.line.y=element_line()) +
  theme(strip.background=element_rect(fill='gray95', colour='transparent'),
        legend.key.size=unit(.6, "lines"), 
        legend.position=c(1, 1.1), legend.justification=c(1, 1) )

ggsave('plots/BZSposter2017/lags_with_iptg.pdf', width=180/25.4/2, height=100/25.4/2)
  
```

