---
title: "Single-cell analysis of the glucose/lactose diauxic shift in E. coli"
author: Thomas Julou
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: false
---

This is a companion website describing the data analysis underlying the article [*Subpopulations of sensorless bacteria drive fitness in fluctuating environments*](https://doi.org/10.1101/2020.01.04.894766).

This project consists for a large part of single-cell experiments done using microfluidics:

- [Lags estimation](MoM_lacInduction_Lags_Estimation.html)
- [Controls](MoM_lacInduction_Controls.html)
- [Induction and memory of the native *lac* operon](MoM_lacInduction_Native.html)
- [Perturbations of the repressed lac operon expression](MoM_lacInduction_PerturbRepressed.html)
- [*lac* induction at lower lactose concentrations](MoM_lacInduction_lowLactose.html)
- [Simulation of population lags from distribution of single-cell lags](MoM_lacInduction_PopLagSimul.html)


In addition, we used an array of other techniques and analysis:

- [Fluorescence lifetime experiments](MoM_lacInduction_FLIM.html)
- [Population lags in diauxie experiments](MoM_lacInduction_diauxieGrowthCurves.html)
- [Identification of SMT candidates in MS data](MoM_lacInduction_QMS.html)


Available material includes:

- raw image datasets, as well as images processed for analysis (https://doi.org/10.17867/10000149),
- data tables obtained from image analysis (https://doi.org/10.5281/zenodo.3894719),
- analysis source code and additional data (https://github.com/julou/MoM_lacInduction).


## Mother Machine experiments

In typical experiments, E. coli is grown in a Dual-Input Mother Machine where glucose and lactose alternate every 4 hours:

<iframe width="853" height="480" src="https://www.youtube.com/embed/2Tznm868fmc?rel=0" frameborder="0" allowfullscreen></iframe>


### Dataset description

```{r}
mygl <- myframes %>% ungroup() %>% 
  filter(!discard_top, !discard_start) %>% 
  mutate(label=rename_conds(condition)) %>% 
  # start from the total number of observations
  group_by(condition, label, date, pos, gl) %>%
  summarise(n_obs=n()) %>%
  # add the number of dividing cells
  left_join(myframes %>%
              ungroup() %>% 
              filter(!discard_top, end_type=='div') %>% 
              group_by(date, pos, gl, id) %>% 
              filter(!any(discard_start)) %>% 
              slice(1) %>% 
              group_by(date, pos, gl) %>% 
              summarise(n_div_cells=n()) ) %>%
  # add the number of frames
  left_join(myframes %>%
              group_by(date, pos, gl) %>% 
              summarise(nframes=max(frame)) )

```


```{r}
(mytables[['expts_list']] <- mygl %>% ungroup() %>% 
   # mutate(condition=factor(condition, levels=unique(condition_acq_times$condition)) ) %>% 
   group_by(condition, label, date) %>%
   summarise(n_lanes=length(unique(interaction(date, pos, gl))),
             # n_div_cells=sum(is.na(n_div_cells)),
             n_div_cells=sum(n_div_cells),
             n_obs=sum(n_obs) ) %>% 
   left_join(
     mycells_switching %>% 
       ungroup %>% 
       filter(! date %in% discarded_dates) %>% 
       group_by(condition, date) %>%
       summarise(
         n_switch=n(),
         n_lags=sum(!is.na(lag_200)),
         n_discard=sum(logl_time_slope_before <= min_growth_rate, na.rm=TRUE) ) %>% 
       mutate(n_discard=ifelse(condition=='switch_gly_lac', NA, n_discard) )
   )
) %>% 
  filter(! date %in% discarded_dates) %>% 
  knitr::kable() %>% 
  kableExtra::kable_styling() %>% 
  identity()

```



## Colophon

Package versions are managed with `renv` and can be found in the corresponding [lockfile](https://github.com/julou/MoM_lacInduction/blob/master/renv.lock).

```{r eval=FALSE, echo=FALSE}
devtools::loaded_packages() %>% 
  mutate(version = unlist(lapply(package, function(x) as.character(packageVersion(x)))) ) %>% 
  arrange(package) %>% 
  select(-path) %>% 
  knitr::kable(row.names=FALSE)

```


