---
title: "*lac* induction and growth lag in naive cells"
author: Thomas Julou
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

## Distribution of lags

For the first switch, all conditions (without IPTG) should be equivalent:
  
```{r}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(logl_time_slope_pre > min_growth_rate) %>% 
  filter(switch_idx==1) %>%
  gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
  mutate(lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         lag=ifelse(is.na(lag), -1, lag)) %>% 
  ggplot(aes(x=lag/60)) +
  stat_ecdf(aes(y=1-..y.., col=variable)) +
  labs(x='lag after the switch (min)', y='reverse cumulative probability', title='distribution of lags at 1st switch (without 8h)') +
  expand_limits(x=0)

mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>%
  filter(logl_time_slope_pre > min_growth_rate) %>% 
  filter(switch_idx==1) %>%
  gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
  mutate(lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         lag=ifelse(is.na(lag), -10, lag)) %>% 
  ggplot(aes(x=lag/60)) +
  facet_grid(variable~.) +
  geom_step_hist(aes(col=factor(switch_idx)), binwidth=3, position='identity') +
  geom_histogram(aes(fill=factor(switch_idx)), binwidth=3, position='identity', 
                 col='transparent', alpha=.2) +
  labs(x='lag after the switch (min)', title='distribution of finite lags at 1st switch (without 8h)', col='switch', fill='switch') +
  xlim(-1, 240) +
  guides(col='none', fill='none')

```

Highlight on the long lag condition: the high proportion of non switching cells initially seen corresponds to cells growing very slowly before the switch (by chance very abundant on this day) 

```{r}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  filter(logl_time_slope_pre > min_growth_rate) %>%
  filter(switch_idx==1) %>%
  gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
  group_by(date) %>% 
  mutate(lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60)) %>% 
  (function(.df)
    ggplot(.df %>% filter(condition!='switch_long_lac'), aes(x=lag/60)) +
     facet_grid(variable~.) +
     stat_ecdf(aes(y=1-..y.., col='04h')) +
     stat_ecdf(aes(y=1-..y.., group=date, col='04h'), alpha=.2, size=.5) +
     stat_ecdf(aes(y=1-..y.., col='24h'),
               data=.df %>% filter(condition=='switch_long_lac')) +
     labs(x='lag after the switch (min)', y='reverse cumulative probability', col='lactose\nduration') +#
     # , title='distribution of lags at 1st switch') +
scale_y_log10() +
          expand_limits(x=0) )

```

```{r BZSposter, eval=FALSE}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  filter(logl_time_slope_pre > min_growth_rate) %>% 
  filter(switch_idx==1) %>%
  gather(variable, lag, growth_lag, gfp_lag, factor_key=TRUE) %>% 
  mutate(variable=forcats::fct_recode(variable, 'growth lag'='growth_lag', 'lac induction lag'='gfp_lag'),
         lag=ifelse(lag>240*60, Inf, lag) ) %>% #, 
         # lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         # lag=ifelse(is.na(lag), -200, lag)) %>% 
  ggplot(aes(x=lag/60)) +
  # facet_grid(variable~.) +
  geom_histogram(aes(fill=variable), binwidth=6, position='identity', 
                 col='transparent', alpha=.1) +
  geom_step_hist(aes(col=variable), binwidth=6, position='identity') +
  labs(x='lag (min)') +
  scale_x_continuous(breaks=seq(0, 200, 50)) +
  theme_classic(base_size=8) +
  theme(axis.line.x=element_line(), axis.line.y=element_line()) +
  theme(legend.position=c(1, 1), legend.justification=c(1, 1) )
# ggsave('Rplot.pdf', width=180/25.4/2, height=100/25.4/2)
  
# # with lag_200 instead
# mycells_switching %>% ungroup %>%
#   filter(!date %in% discarded_dates) %>%
#   filter(switch_idx==1) %>%
#   filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
#   filter(logl_time_slope_pre > min_growth_rate) %>% 
#   gather(variable, lag, growth_lag, lag_200, factor_key=TRUE) %>%
#   mutate(variable=forcats::fct_recode(variable, 'growth lag'='growth_lag', 'lac induction lag'='lag_200'),
#          lag=ifelse(lag>240*60, Inf, lag) ) %>% #,
#          # lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
#          # lag=ifelse(is.na(lag), -200, lag)) %>%
#   ggplot(aes(x=lag/60)) +
#   facet_grid(variable~.) +
#   geom_histogram(aes(fill=variable), binwidth=6, position='identity',
#                  col='transparent', alpha=.1) +
#   geom_step_hist(aes(col=variable), binwidth=6, position='identity') +
#   labs(x='lag (min)') +
#   scale_x_continuous(breaks=seq(0, 200, 50)) +
#   theme_classic() +
#   theme(axis.line.x=element_line(), axis.line.y=element_line()) +
#   theme(legend.position=c(1, 1), legend.justification=c(1, 1) )

```

check for a possible day effect:

```{r}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  filter(logl_time_slope_pre > min_growth_rate) %>% 
  filter(switch_idx==1) %>%
  gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
  mutate(lag=ifelse(is.finite(lag) & lag>240*60, 240*60, lag), 
         lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         lag=ifelse(is.na(lag), -10, lag)) %>% 
  filter(variable=='lag_200') %>% 
  ggplot(aes(x=lag/60)) +
  facet_grid(date+condition~.) +
  geom_step_hist(aes(col=factor(switch_idx)), binwidth=6, position='identity') +
  geom_histogram(aes(fill=factor(switch_idx)), binwidth=6, position='identity', 
                 col='transparent', alpha=.2) +
  labs(x='induction lag (min)', title='naive cells', col='switch', fill='switch') +
  xlim(-1, 250) +
  guides(col='none', fill='none')

mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(!condition %in% c('switch_gly_lac', 'switch_lac_IPTG500uM')) %>% 
  filter(logl_time_slope_pre > min_growth_rate) %>% 
  filter(switch_idx==1) %>%
  gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
  mutate(lag=ifelse(is.finite(lag) & lag>240*60, 240*60, lag), 
         lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         lag=ifelse(is.na(lag), -10, lag)) %>% 
  filter(variable=='lag_200') %>% 
  ggplot(aes(x=lag/60)) +
  facet_wrap(~date+condition, dir='v', ncol=3, scales='free_y', strip.position='right') +
  geom_step_hist(aes(col=factor(switch_idx)), binwidth=6, position='identity') +
  geom_histogram(aes(fill=factor(switch_idx)), binwidth=6, position='identity', 
                 col='transparent', alpha=.2) +
  labs(x='induction lag (min)', title='naive cells', col='switch', fill='switch') +
  xlim(-1, 250) +
  guides(col='none', fill='none')

```

how the bimodal lag of GFP induction becomes more unimodal for growth is not clear: it is tempting to hypothesize that there is some additional variability in the time different cells need to restart growing after expressing GFP / that they induce GFP at different rates, hence bluring the 2 peaks into a single broader peak.

```{r}
mycells_switching %>% 
  mutate(lag_200=as.character(lag_200), lag_200=ifelse(is.na(lag_200), 'NA', lag_200),
         lag_200=ifelse(is.finite(as.numeric(lag_200)) & as.numeric(lag_200)<50*60, '<50', lag_200),
         lag_200=ifelse(is.finite(as.numeric(lag_200)) & as.numeric(lag_200)>=50*60, '≥50', lag_200) ) %>%
  group_by(switch_idx, lag_200) %>% 
  summarise(n=n()) %>% 
  group_by(switch_idx) %>% mutate(p=n/sum(n)) %>% 
  knitr::kable(digits=3)

```

Check the fraction of arrested cells in mother cells since they have no bias of long lags being more truncated:

```{r}
mycells_switching %>% 
  filter(cell_num_from_bottom==0) %>% # keep only mother cells
  mutate(lag_200=as.character(lag_200), lag_200=ifelse(is.na(lag_200), 'NA', lag_200),
         lag_200=ifelse(is.finite(as.numeric(lag_200)) & as.numeric(lag_200)<50*60, '<50', lag_200),
         lag_200=ifelse(is.finite(as.numeric(lag_200)) & as.numeric(lag_200)>=50*60, '≥50', lag_200) ) %>%
  group_by(switch_idx, lag_200) %>% 
  summarise(n=n()) %>% 
  group_by(switch_idx) %>% mutate(p=n/sum(n)) %>% 
  knitr::kable(digits=3)

```



## Lag heritability

Let's look at the heritability of the lag in naive cells, i.e. how much a common history (and hence cellular state) determines the duration of the lag.
first visualize the pairwise correlations to understand the heritability estimator.

focus on first switch with induction lag:

```{r}
mypairs_switching %>% 
  filter(rel!='niece') %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  filter(logl_time_slope_pre_1 > min_growth_rate, logl_time_slope_pre_2 > min_growth_rate) %>% 
  filter(switch_idx==1, variable=='lag_200') %>%
  # create a symmetric scatter by duplicating
  group_by(condition, date, rel, variable) %>% 
  do((function(.df) bind_rows(.df, rename(.df, value_1=value_2, value_2=value_1)) )(.)) %>% 
  mutate(value_1=ifelse(value_1<240*60, value_1, Inf), value_2=ifelse(value_2<240*60, value_2, Inf)) %>% 
  # filter(is.finite(value_1), is.finite(value_2)) %>% 
  ggplot(aes(value_1/60, value_2/60, col=rel)) +
# ggplot(filter(mypairs_switching, rel!='niece'), aes(lag_200_1/60, lag_200_2/60, col=rel)) +
  facet_grid(date+condition~rel) +
  geom_abline(col='gray50') +
  geom_point(alpha=.5, size=1) +
  expand_limits(x=0, y=0) +
  labs(x='induction lag cell 1 (min)', y='induction lag cell 2 (min)', col='relationship') +
  # scale_colour_manual(values=c('sisters'=brewer_cols[1], 'cousins2'=brewer_cols[3])) +
  guides(col='none') +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  coord_equal()

mypairs_switching %>% 
  filter(rel!='niece') %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  filter(logl_time_slope_pre_1 > min_growth_rate, logl_time_slope_pre_2 > min_growth_rate) %>% 
  filter(switch_idx==1, variable=='lag_200') %>%
  mutate(value_1=ifelse(value_1<240*60, value_1, Inf), value_2=ifelse(value_2<240*60, value_2, Inf)) %>% 
  mutate(both_finite=is.finite(value_1) & is.finite(value_2)) %>%
  group_by(condition, date, variable, rel) %>% 
  summarise(pearson=cor_sym(value_1[both_finite], value_2[both_finite], method="pearson")^2, 
            spearman=cor_sym(value_1, value_2, method="spearman", use="complete.obs")^2, 
            n=n(), n_finite=sum(both_finite)) %>% 
  gather(type, r2, pearson, spearman) %>% 
  # filter(type=='pearson') %>% 
  (function(.df) ggplot(.df, aes(interaction(condition, date, sep='\n'), r2)) +
     facet_grid(type~., labeller=as_labeller(switch_facets_labels)) +
     geom_bar(aes(fill=rel), position='dodge', stat='identity', show.legend=FALSE) +
     # geom_text(aes(y=0.99, label=paste0("n = ", n_finite)), angle=90, hjust=1) +
     expand_limits(y=1) + labs(y='squared correlation coeff') +
     theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1), axis.title.x=element_blank())
  )

```


```{r}
mypairs_switching %>% 
  filter(rel %in% c('sisters', 'cousins2')) %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  filter(logl_time_slope_pre_1 > min_growth_rate, logl_time_slope_pre_2 > min_growth_rate) %>% 
  filter(switch_idx==1) %>%
  # create a symmetric scatter by duplicating
  group_by(condition, date, rel, variable) %>% 
  do((function(.df) bind_rows(.df, rename(.df, value_1=value_2, value_2=value_1)) )(.)) %>% 
  mutate(value_1=ifelse(value_1<240*60, value_1, Inf), value_2=ifelse(value_2<240*60, value_2, Inf)) %>% 
  # filter(is.finite(value_1), is.finite(value_2)) %>% 
  ggplot(aes(value_1/60, value_2/60, col=rel)) +
# ggplot(filter(mypairs_switching, rel!='niece'), aes(lag_200_1/60, lag_200_2/60, col=rel)) +
  # facet_wrap(~switch_idx, scales='free', ncol = 3) +
     facet_grid(rel~variable) +
  geom_abline(col='gray50') +
  geom_point(alpha=.2, size=.5) +
  expand_limits(x=0, y=0) +
  labs(x='induction lag cell 1 (min)', y='induction lag cell 2 (min)', col='relationship') +
  scale_colour_manual(values=c('sisters'=brewer_cols[1], 'cousins2'=brewer_cols[3])) +
  guides(col='none') +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) +
  coord_equal()

```

```{r}
mypairs_switching %>% 
  filter(rel!='niece') %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  filter(logl_time_slope_pre_1 > min_growth_rate, logl_time_slope_pre_2 > min_growth_rate) %>% 
  filter(switch_idx==1) %>%
  mutate(value_1=ifelse(value_1<240*60, value_1, Inf), value_2=ifelse(value_2<240*60, value_2, Inf)) %>% 
  mutate(both_finite=is.finite(value_1) & is.finite(value_2)) %>%
  group_by(variable, rel) %>% 
  summarise(pearson=cor_sym(value_1[both_finite], value_2[both_finite], method="pearson")^2, 
            spearman=cor_sym(value_1, value_2, method="spearman", use="complete.obs")^2, 
            n=n(), n_finite=sum(both_finite)) %>% 
  gather(type, r2, pearson, spearman) %>% 
  # filter(type=='pearson', variable=='lag_200') %>% 
  # filter(type=='pearson') %>% 
  (function(.df) ggplot(.df, aes(rel, r2)) +
     facet_grid(type~variable, labeller=as_labeller(switch_facets_labels)) +
     geom_bar(aes(fill=rel), stat='identity', show.legend=FALSE) +
     geom_text(aes(y=0.99, label=paste0("n = ", n_finite)), angle=90, hjust=1) +
     expand_limits(y=1) + labs(x='relationship', y='squared pearson correlation r2') +
     theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1))
  )

```

By comparison, let's look at the heritability of various traits in constant environment:

```{r}
mypairs_constant %>% 
  filter(div_min==div_max) %>% 
  group_by(condition, variable, rel) %>% 
  summarise(r2=cor(value_1,  value_2, use="complete.obs", method="pearson")^2, n=n()) %>% 
  ggplot(aes(variable, r2)) +
  facet_wrap(~condition) +
  geom_bar(aes(fill=rel), stat='identity', pos='dodge') +
  geom_text(aes(y=1, label=sprintf("(n = %d)", n), group=rel), 
            angle=90, hjust=1, pos=position_dodge(width=0.8)) +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        legend.position='top')

bind_rows(
  mypairs_constant %>% 
    filter(div_min==div_max) %>% 
    filter(variable!='dt') %>% #, !(condition=='glucose' & variable %in% c('c_birth', 'c_div'))) %>% 
    group_by(condition, variable, rel) %>% 
    summarise(r2=cor_sym(value_1,  value_2, use="complete.obs", method="pearson")^2, n=n()),
  mypairs_switching %>% 
    filter(!date %in% discarded_dates) %>%
    filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
    filter(logl_time_slope_pre_1 > min_growth_rate, logl_time_slope_pre_2 > min_growth_rate) %>% 
    filter(rel!='niece', switch_idx==1) %>% 
    mutate(value_1=ifelse(value_1<240*60, value_1, Inf), value_2=ifelse(value_2<240*60, value_2, Inf)) %>% 
    mutate(lag_200_1=value_1, lag_200_2=value_2) %>% 
    filter(!is.infinite(lag_200_1), !is.infinite(lag_200_2)) %>% 
    group_by(rel) %>% 
    summarise(r2=cor_sym(lag_200_1, lag_200_2, use="complete.obs", method="pearson")^2, n=n()) %>% 
    mutate(condition='switch', variable='lag')
) %>% 
  mutate(rel=forcats::fct_recode(rel, sisters="sisters", "1st cousins"="cousins1", "2nd cousins"="cousins2") %>% (forcats::fct_drop)) %>% #,
  ggplot(aes(variable, r2)) +
  facet_grid(.~condition, scales="free_x", space = "free_x") +
  geom_bar(aes(fill=rel), stat='identity', pos='dodge') +
  geom_text(aes(y=1.2, label=sprintf("(n = %d)", n), group=rel), 
            angle=90, hjust=0.9, pos=position_dodge(width=0.8)) +
  scale_y_continuous(breaks = seq(0, 1, .5)) +
  labs(y='Pearson correlation coefficient r2', fill='relationship') +
  theme_classic() + theme(axis.line.x=element_line(), axis.line.y=element_line()) +
  theme(#legend.position=c(0.01, 1), legend.justification=c(0, 1),
        panel.border=element_rect(size=.3, fill='transparent')) +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1)) #, legend.position='top') 

```


```{r BZSposter, eval=FALSE}
bind_rows(
  mypairs_constant %>% 
    filter(div_min==div_max) %>% 
    filter(condition=='lactose', variable %in% c('alpha', 'l_birth', 'c_birth')) %>% 
    group_by(condition, variable, rel) %>% 
    summarise(r2=cor(value_1,  value_2, use="complete.obs", method="pearson")^2, n=n()),
  #
  mypairs_switching %>% 
    filter(!date %in% discarded_dates) %>%
    filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
    filter(logl_time_slope_pre_1 > min_growth_rate, logl_time_slope_pre_2 > min_growth_rate) %>% 
    filter(rel!='niece', switch_idx==1, variable %in% c('gfp_lag', 'growth_lag')) %>% 
    mutate(value_1=ifelse(value_1<240*60, value_1, Inf), value_2=ifelse(value_2<240*60, value_2, Inf)) %>% 
    mutate(both_finite=is.finite(value_1) & is.finite(value_2)) %>%
    group_by(variable, rel) %>% 
    summarise(r2=cor(value_1[both_finite], value_2[both_finite], method="pearson")^2, 
              n=sum(both_finite))
) %>% 
  ungroup %>% mutate(variable=factor(variable, 
                                     levels=c('gfp_lag', 'growth_lag', 'alpha', 'l_birth', 'c_birth')) %>% 
                       forcats::fct_recode('lac induction\nlag'='gfp_lag', 'growth lag'='growth_lag', 'growth rate'='alpha', 'size at birth'='l_birth', '[GFP] at birth'='c_birth') ) %>% 
  ggplot(aes(variable, r2)) +
  geom_bar(aes(fill=rel), stat='identity', pos='dodge') +
  geom_text(aes(y=1, label=sprintf("(n = %d)", n), group=rel), 
            angle=90, hjust=1, vjust=.0, pos=position_dodge(width=0.8), size=2) +
  scale_x_discrete() +
  labs(y='Pearson correlation coeff r2', fill='relationship') +
  theme_classic(base_size=8) + 
  theme(axis.line.x=element_line(), axis.line.y=element_line()) +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        axis.title.x=element_blank())
ggsave('Rplot.pdf', width=180/25.4/2, height=100/25.4/2)

```

We want to understand (i) the molecular processes underlying the multimodal distribution of the lags, and (ii) what are theounds shared by cellullarcompurelated cells that determines their lags. Actually these questions might very well be linked. 


## Lag heritability vs cell to cell communication

Focus on the first switch.

```{r}
mypairs_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  filter(logl_time_slope_pre_1 > min_growth_rate, logl_time_slope_pre_2 > min_growth_rate) %>% 
  filter(rel!='niece', switch_idx==1, variable %in% c('lag_200')) %>% 
  # filter(!condition %in% c('switch_lac_IPTG500uM', 'switch_long_lac', 'switch_long_lac_hiExpr')) %>% 
  # mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$'), 'switch', condition)) %>%
  mutate(cond=interaction(condition, date, drop = T),
         pos_dist=abs(cell_num_1-cell_num_2),
         rel=interaction(div_min, div_max, drop=TRUE)) %>% 
  (function(.df) bind_rows(.df, rename(.df, value_1=value_2, value_2=value_1))) %>% 
  filter(pos_dist<7, rel!='1.3') %>% 
  ggplot() +
  facet_grid(rel~pos_dist) +
  geom_abline(size=.2, col=brewer_cols[2]) +
  geom_vline(xintercept=55, lty='dotted', size=.2) +
  geom_hline(yintercept=55, lty='dotted', size=.2) +
  geom_point(aes(value_1/60, value_2/60), alpha=.2, size=.5)

mypairs_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  filter(logl_time_slope_pre_1 > min_growth_rate, logl_time_slope_pre_2 > min_growth_rate) %>% 
  filter(rel!='niece', switch_idx==1, variable %in% c('lag_200')) %>% 
  # filter(!condition %in% c('switch_lac_IPTG500uM', 'switch_long_lac', 'switch_long_lac_hiExpr')) %>% 
  # mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$'), 'switch', condition)) %>%
  mutate(cond=interaction(condition, date, drop = T),
         pos_dist=abs(cell_num_1-cell_num_2),
         rel=interaction(div_min, div_max, drop=TRUE)) %>% 
  (function(.df) bind_rows(.df, rename(.df, value_1=value_2, value_2=value_1))) %>% 
  filter(pos_dist<7, rel!='1.3') %>% 
  group_by(rel, pos_dist) %>% 
  select(rel, pos_dist, value_1, value_2) %>% 
  filter(is.finite(value_1), is.finite(value_2)) %>% 
  summarise(r2=cor(value_1, value_2)^2, n=n()) %>% 
  filter(n>10) %>% 
  # mutate(r2=-log(1-r2)) %>%
  ggplot(aes(pos_dist, forcats::fct_rev(rel))) +
  geom_tile(aes(fill=r2, alpha=n)) +
  geom_text(aes(label=round(r2, digits=2), size=n)) +
  labs(y='relationship', title="Pearson r2") +
  # labs(title="-log(1-r2)") +
  scale_x_continuous(breaks=0:10) +
  theme_classic() +
  theme(axis.line = element_blank(), axis.ticks = element_blank())
  # select(-n) %>% 
  # spread(pos_dist, r2) %>% 
  # knitr::kable(digits=2)
  
```


```{r}
mypairs_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  filter(logl_time_slope_pre_1 > min_growth_rate, logl_time_slope_pre_2 > min_growth_rate) %>% 
  filter(rel!='niece', switch_idx==1, variable %in% c('lag_200')) %>% 
  # filter(!condition %in% c('switch_lac_IPTG500uM', 'switch_long_lac', 'switch_long_lac_hiExpr')) %>% 
  # mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$'), 'switch', condition)) %>%
  mutate(cond=interaction(condition, date, drop = T),
         pos_dist=abs(cell_num_1-cell_num_2),
         rel=interaction(div_min, div_max, drop=TRUE)) %>% 
  (function(.df) bind_rows(.df, rename(.df, value_1=value_2, value_2=value_1))) %>% 
  filter(pos_dist<7, rel!='1.3') %>% 
  group_by(condition, date, rel, pos_dist) %>% 
  select(rel, pos_dist, value_1, value_2) %>% 
  filter(is.finite(value_1), is.finite(value_2)) %>% 
  summarise(r2=cor(value_1, value_2)^2, n=n()) %>% 
  filter(n>10, pos_dist<5) %>% 
  # mutate(r2=-log(1-r2)) %>%
  ggplot(aes(pos_dist, forcats::fct_rev(rel))) +
  facet_wrap(~date+condition, dir='v', ncol=2, scales='free_y', strip.position='right') +
  geom_tile(aes(fill=r2, alpha=n)) +
  geom_text(aes(label=round(r2, digits=2), size=n)) +
  labs(y='relationship', title="Pearson r2") +
  # labs(title="-log(1-r2)") +
  scale_x_continuous(breaks=0:10) +
  # scale_fill_gradient(name='r', low=brewer_cols[2], high=brewer_cols[1], na.value="grey95") +
  theme_classic() +
  theme(axis.line = element_blank(), axis.ticks = element_blank())
  # select(-n) %>% 
  # spread(pos_dist, r2) %>% 
  # knitr::kable(digits=2)

mypairs_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  filter(logl_time_slope_pre_1 > min_growth_rate, logl_time_slope_pre_2 > min_growth_rate) %>% 
  filter(rel!='niece', switch_idx==1, variable %in% c('lag_200')) %>% 
  # filter(!condition %in% c('switch_lac_IPTG500uM', 'switch_long_lac', 'switch_long_lac_hiExpr')) %>% 
  # mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$'), 'switch', condition)) %>%
  mutate(cond=interaction(condition, date, drop = T),
         pos_dist=abs(cell_num_1-cell_num_2),
         rel=interaction(div_min, div_max, drop=TRUE)) %>% 
  (function(.df) bind_rows(.df, rename(.df, value_1=value_2, value_2=value_1))) %>% 
  filter(pos_dist<5, rel!='1.3') %>% 
  ggplot() +
  facet_grid(condition+date+rel~pos_dist) +
  geom_abline(size=.2, col=brewer_cols[2]) +
  geom_vline(xintercept=55, lty='dotted', size=.2) +
  geom_hline(yintercept=55, lty='dotted', size=.2) +
  geom_point(aes(value_1/60, value_2/60), alpha=.5, size=1)


```

Let's check whether all cells are induced after the first one in a GL:
- if the lag is dominated by communication, lag delay should be short and shouldnt depend on genealogy
- is the lag is dominated by genealogy, the lag delay should be shorter for closely related cells 

```{r}
mypairs_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  filter(logl_time_slope_pre_1 > min_growth_rate, logl_time_slope_pre_2 > min_growth_rate) %>% 
  filter(switch_idx==1, rel!='niece') %>% 
  mutate(value_min=ifelse(value_1<value_2, value_1, value_2),
         value_max=ifelse(value_1>=value_2, value_1, value_2)) %>% 
  group_by(date, pos, gl) %>% 
  filter(value_min==min(value_min)) %>% 
  mutate(cond=interaction(condition, date, drop = T),
         pos_dist=abs(cell_num_1-cell_num_2),
         rel=interaction(div_min, div_max, drop=TRUE)) %>% 
  filter(pos_dist<7, rel!='1.3') %>% 
  ggplot() +
  facet_grid(rel~.) +
  geom_histogram(aes((value_max-value_min)/60), binwidth=3) +
  labs(x='lag delay (min)')

mypairs_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac') %>%
  filter(logl_time_slope_pre_1 > min_growth_rate, logl_time_slope_pre_2 > min_growth_rate) %>% 
  filter(switch_idx==1, rel!='niece') %>% 
  mutate(value_min=ifelse(value_1<value_2, value_1, value_2),
         value_max=ifelse(value_1>=value_2, value_1, value_2)) %>% 
  group_by(date, pos, gl) %>% 
  filter(value_min==min(value_min)) %>% 
  mutate(cond=interaction(condition, date, drop = T),
         pos_dist=abs(cell_num_1-cell_num_2),
         rel=interaction(div_min, div_max, drop=TRUE)) %>% 
  filter(pos_dist<7, rel!='1.3') %>% 
  ggplot() +
  facet_grid(rel~.) +
  # geom_histogram(aes((value_max-value_min)/60), binwidth=3) +
  # labs(x='lag delay (min)')
  geom_hline(yintercept=55, lty='dotted', size=.2) +
  geom_point(aes((value_max-value_min)/60, value_min/60), alpha=.2, size=.5) +
  labs(x='lag delay (min)', y='shortest lag in GL')
  
```




## Effect of spontaneous *lac* expression

hypothesized to explain how the positive feedback kicks off.
can be manipulated by titrating a given fraction of LacI with IPTG.

```{r}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(logl_time_slope_pre > min_growth_rate) %>% 
  filter(switch_idx==1) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac' | 
           str_detect(condition, 'withIPTG') | str_detect(condition, 'lacIoe')) %>% 
  mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac',
         'lactose', condition)) %>%
  gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
  mutate(lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         lag=ifelse(is.na(lag), -1, lag)) %>% 
  ggplot(aes(x=lag/60)) +
  facet_grid(variable~.) +
  stat_ecdf(aes(y=1-..y.., col=condition)) +
  labs(x='lag after the switch (min)', y='reverse cumulative probability', title='distribution of lags at 1st switch') +
  expand_limits(x=0) + 
  xlim(0, 240)

mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_datesss) %>%
  filter(logl_time_slope_pre > min_growth_rate) %>% 
  filter(switch_idx==1) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac' | 
           str_detect(condition, 'withIPTG') | str_detect(condition, 'lacIoe')) %>% 
  mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac', 'lactose', condition)) %>%
  gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
  mutate(lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         lag=ifelse(is.na(lag), -200, lag)) %>% 
  ggplot(aes(x=condition, y=lag/60)) +
  facet_grid(variable~.) +
  geom_violin(aes(group=date, fill=condition), col='transparent', alpha=.4, position='identity') +
  labs(y='lag after the switch (min)') +
  ylim(0, 240)

```


Let's check that the *lac* promoter is not induced before the switch:

```{r}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(logl_time_slope_pre > min_growth_rate) %>% 
  filter(switch_idx==1) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac' | 
           str_detect(condition, 'withIPTG') | str_detect(condition, 'lacIoe')) %>% 
  mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac', 'no IPTG', condition)) %>%
  gather(variable, lag, growth_lag, gfp_lag, lag_200, factor_key=TRUE) %>% 
  mutate(lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         lag=ifelse(is.na(lag), -200, lag)) %>% 
  ggplot(aes(condition, gfp_ini)) +
  geom_violin()
    # labs(x='lag after the switch (min)', title='distribution of finite lags at 1st switch') +
  # xlim(0, 240)

```


```{r presentation, eval=FALSE}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_datesss) %>%
  filter(logl_time_slope_pre > min_growth_rate) %>% 
  filter(switch_idx==1) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac' | 
           str_detect(condition, 'withIPTG') | str_detect(condition, 'lacIoe')) %>% 
  mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac', 'control', condition),
         condition=forcats::fct_relevel(condition, 'switch_withIPTG5uM', 'switch_withIPTG1uM', "control", 
                                        'switch_lacIoe_withIPTG10uM', 'switch_lacIoe'),
         treatment=forcats::fct_recode(condition, 'less LacI'='switch_withIPTG1uM', 'less LacI'='switch_withIPTG5uM',
                                       'more LacI'='switch_lacIoe', 'more LacI'='switch_lacIoe_withIPTG10uM'),
         condition=forcats::fct_recode(condition, 'with 1µM IPTG'='switch_withIPTG1uM', 'with 5µM IPTG'='switch_withIPTG5uM',
                                       'with LacI plasmid'='switch_lacIoe', 
                                       'with LacI plasmid\n+ 10µM IPTG'='switch_lacIoe_withIPTG10uM')) %>% 
  # mutate(lag=lag_200) %>% 
  gather(variable, lag, growth_lag, gfp_lag, factor_key=TRUE) %>% 
  mutate(variable=forcats::fct_recode(variable, 'growth lag'='growth_lag', 'lac induction lag'='gfp_lag'),
         lag=ifelse(lag>240*60, Inf, lag) ) %>% #, 
  mutate(lag=ifelse(!is.infinite(lag), lag, max(lag[is.finite(lag)])+60),
         lag=ifelse(is.na(lag), -200, lag)) %>% 
  ggplot(aes(x=condition, y=lag/60)) +
  facet_grid(variable~.) +
  geom_hline(yintercept=48, lty='dashed') +
  geom_violin(aes(group=date, fill=treatment, col=treatment), alpha=.4, position='identity') +
  labs(y='lag (min)') +
  scale_color_manual(values=brewer_cols[c(2, 1, 3)], guide='none') +
  scale_fill_manual(values=brewer_cols[c(2, 1, 3)], guide='none') +
  ylim(0, 240) +
  theme_classic(base_size=8) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_text(angle=20, hjust=0.8, vjust=0.9),
        strip.background=element_rect(fill='gray95', colour='transparent'), ) 
# ggsave('Rplot.pdf', width=180/25.4/2, height=100/25.4/2)

```

Visualize fluctuations of LacZ-GFP in the glucose phase

```{r}
ggplot(data=NULL, aes(time_sec, gfp_nb, col=condition, group=cell)) +
  geom_line(data=myframes %>% filter(!discard_start, !discard_top) %>% filter(condition %in% c('switch_16h'), medium=='glucose', m_cycle==1), alpha=.2) + #'switch_04h',
  geom_line(data=myframes %>% filter(!discard_start, !discard_top) %>% filter(condition=='switch_withIPTG5uM', medium=='glucose+IPTG', m_cycle==1), alpha=.2)

```

## Effect of CRP regulation

distribution of lags for naive cells grown in glucose (CRP active, no inducer exclusion):

```{r}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(switch_idx==1) %>%
  filter(condition!='switch_lac_IPTG500uM') %>% 
  mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac', 'switch_glc_lac', condition),
         condition=forcats::fct_recode(condition, 'with 1µM IPTG'='switch_withIPTG1uM',
                                       'with 5µM IPTG'='switch_withIPTG5uM') ) %>%
  filter(condition %in% c('switch_glc_lac', 'switch_gly_lac')) %>% 
  gather(variable, lag, growth_lag, gfp_lag, factor_key=TRUE) %>% 
  mutate(variable=forcats::fct_recode(variable, 'growth lag'='growth_lag', 'lac induction lag'='gfp_lag')) %>% 
  mutate(lag=ifelse(lag>240*60, Inf, lag),
         lag=ifelse(!is.infinite(lag), lag, 241*60),
         lag=ifelse(is.na(lag), -200, lag)) %>% 
  ggplot(aes(x=lag/60)) +
  facet_grid(variable~., scales='free_y') +
  geom_histogram(aes(y=..density.., fill=condition), binwidth=6, position='identity', 
                 col='transparent', alpha=.2) +
  geom_step_hist(aes(y=..density.., col=condition), binwidth=6, position='identity') +
  labs(x='lag (min)') +
  xlim(0, 240) +
  theme_classic() + 
  theme(axis.line.x=element_line(), axis.line.y=element_line()) +
  theme(strip.background=element_rect(fill='gray95', colour='transparent'),
        legend.key.size=unit(.8, "lines"), 
        legend.position=c(1, 1), legend.justification=c(1, 1) )

```


## Effect of the ramp dynamics

in order to have a not-too-short period with low glucose (> 0) and lactose, cells are switched from 0.002% glucose (≈100µM) to 0.2% lactose. CRP is supposed to kick in around 20-50µM (cf T. Hwa)

first, let's verify that cells grow exponentially and at full speed in 0.002% glucose:

```{r}
mycells_naive <- myframes %>% 
  filter(!discard_top) %>%
  filter(m_start==0) %>% 
  # group_by(condition, date, pos, gl, id, cell, parent_id, genealogy) %>%
  partition(condition, date, pos, gl, id, cell, parent_id, genealogy, cluster=mycluster) %>%
  filter(!any(discard_start), end_type=='div', end_time>m_end) %>% # full cell cycles only
  filter(n()>4) %>% # at least 4 time points
  do((function(.df) {
    # browser()
    .mod_ll_t <- lm( log(length_um)~time_sec, .df)  # use fastLm() for predict
    .time_birth <- first(.df$time_sec)
    .time_div <- last(.df$time_sec)
    .logl <- predict(.mod_ll_t, se.fit=TRUE)
    data.frame(npoints=.mod_ll_t$df.residual+1,
               time_birth=.time_birth, time_div=.time_div, 
               cell_num_from_top=mean(.df$cell_num_in_lane),
               cell_num_from_bottom=mean(.df$total_cell_in_lane-.df$cell_num_in_lane), 
               loglength_start=first(.logl$fit), loglength_startse=first(.logl$se.fit), 
               loglength_end=last(.logl$fit), loglength_endse=last(.logl$se.fit), 
               logl_time_slope=.mod_ll_t$coefficients[2], logl_time_slopesd=summary(.mod_ll_t)$coefficients[2,2], 
               # logl_time_slope=.mod_ll_t$coefficients[2], logl_time_slopesd=.mod_ll_t$stderr[2], 
               logl_time_r2=cor(.df$time_sec, log(.df$length_um))^2 )
  })(.) ) %>% 
  collect() %>% 
  arrange(condition, date, pos, gl, id) %>% 
  mutate(gl_id=gsub('\\.[0-9]+$', '', cell))

mycells_naive %>% 
  filter(!date %in% discarded_date) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition %in% c('switch_gly_lac', 'switch_ramp15min', 'switch_long_lac')) %>% 
  mutate(condition2=ifelse(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac', 'switch_glc_lac', condition)) %>% 
  # filter(!str_detect(condition, 'switch_\\d+h')) %>% 
  ggplot() +
  # facet_grid(condition~.) +
  stat_ecdf(aes(logl_time_r2, col=condition2, group=condition))

mycells_naive %>% ungroup %>%
  filter(!date %in% discarded_date) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition %in% c('switch_gly_lac', 'switch_ramp15min', 'switch_long_lac')) %>% 
  mutate(condition2=ifelse(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac', 'switch_glc_lac', condition),
         condition2=forcats::fct_relevel(factor(condition2), "switch_gly_lac", after=Inf)) %>% 
  # filter(!str_detect(condition, 'switch_\\d+h')) %>% 
  ggplot() +
  # facet_grid(condition~.) +
  stat_ecdf(aes(logl_time_slope*3600/log(2), col=condition2, group=condition)) +
  labs(col="condition", x='growth rate (/h)', y='cumulative density')

```


```{r}
mycells_switching %>% ungroup %>% 
  filter(!date %in% discarded_dates) %>%
  filter(switch_idx==1) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$') | condition %in% c('switch_gly_lac', 'switch_ramp15min', 'switch_long_lac')) %>% 
  mutate(condition=ifelse(str_detect(condition, '^switch_[0-9]+h$') | condition=='switch_long_lac', 'switch_glc_lac', condition),
         condition=forcats::fct_relevel(condition, "switch_gly_lac", after=Inf)) %>% 
  gather(variable, lag, growth_lag, gfp_lag, factor_key=TRUE) %>% 
  mutate(variable=forcats::fct_recode(variable, 'growth lag'='growth_lag', 'lac induction lag'='gfp_lag')) %>% 
  mutate(lag=ifelse(lag>240*60, Inf, lag),
         lag=ifelse(!is.infinite(lag), lag, 241*60),
         lag=ifelse(is.na(lag), -200, lag)) %>% 
  ggplot(aes(x=lag/60)) +
  facet_grid(variable~., scales='free_y') +
  geom_histogram(aes(y=..density.., fill=condition), binwidth=6, position='identity', 
                 col='transparent', alpha=.2) +
  geom_step_hist(aes(y=..density.., col=condition), binwidth=6, position='identity') +
  labs(x='lag (min)') +
  xlim(0, 240) +
  theme_classic() +  # base_size=8
  theme(axis.line.x=element_line(), axis.line.y=element_line()) +
  theme(strip.background=element_rect(fill='gray95', colour='transparent'),
        legend.key.size=unit(.8, "lines"), 
        legend.position=c(1, 1), legend.justification=c(1, 1) )

```


## Relation between bulk and single-cell

```{r}
myframes_switching %>% ungroup %>% 
  filter(!date %in% discarded_date) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>% 
  filter(switch_idx==1) %>% 
  ggplot(aes(time_sec-t_lac_switch, length_um)) +
  geom_line(aes(col=ugen), alpha=.2) +
  scale_colour_periodic_brewer(guide='none') +
  scale_x_hours(0.25) +
  coord_cartesian(xlim=c(-3800, 3800), ylim=c(0, 8))


myframes_switching %>% ungroup %>% 
  filter(!date %in% discarded_date) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>% 
  filter(switch_idx==1) %>% 
  filter(between(time_sec-t_lac_switch, -3600, 3600)) %>% 
  mutate(time_after=time_sec-t_lac_switch) %>% 
  ggplot(aes(time_after, length_um)) +
  geom_line(aes(col=ugen), alpha=.2) +
  geom_line(size=1, data=. %>% group_by(time_after) %>% 
              summarise(n=n(), length_sum=sum(length_um), length_um=length_sum/n) ) +
  scale_colour_periodic_brewer(guide='none') +
  scale_x_hours(0.25) +
  coord_cartesian(ylim=c(0, 4))

myframes_switching %>% ungroup %>% 
  filter(!date %in% discarded_date) %>%
  filter(str_detect(condition, '^switch_[0-9]+h$')) %>% 
  filter(switch_idx==1) %>% 
  filter(between(time_sec-t_lac_switch, -3600, 3600)) %>% 
  mutate(time_after=time_sec-t_lac_switch) %>% 
  group_by(time_after) %>% 
  summarise(n=n(), length_mean=mean(length_um), length_sd=sd(length_um)) %>% 
  ggplot(aes(time_after/60, length_mean)) +
  geom_ribbon(aes(ymin=length_mean-length_sd/sqrt(n), ymax=length_mean+length_sd/sqrt(n)), fill="grey70") +
  geom_line() +
  geom_point(size=.5) +
  coord_cartesian(ylim=c(2, 2.8)) +
  labs(x="time after the switch (min)", y="average length (µm)")


```

